---
title: "Image retrieval using Google Landmarks data"
output: html_notebook
---

Image retrieval using Google Landmarks data - Traditional Models
* Task 1: Import Libraries
* Task 2: Import Images
* Task 3: Data Exploration
* Task 4: Preprocess the Data
* Task 5: Build the Model
* Task 6: Compile the Model
* Task 7: Train and Evaluate the Model
* Task 8: Make Predictions on Test Data


```{r}
## Task 1: Import Libraries ##
EnsurePackage <- function(x){x <- as.character(x)
    if (!require(x,character.only = TRUE)){
      install.packages(pkgs=x, repos = "http://cran.us.r-project.org")
      require(x, character.only = TRUE)
    }
  }
# usage example, to load the necessary library for further processing...

#Data Processing
EnsurePackage("knitr")
EnsurePackage("tidyr")
EnsurePackage("pbapply")
EnsurePackage("sqldf")
EnsurePackage("dplyr")
EnsurePackage("stringr")
EnsurePackage("tidyverse")

#Image Processing
# BiocManager::install("EBImage")
EnsurePackage("EBImage")
EnsurePackage("BiocManager")

#Visualization
EnsurePackage("ggplot2")
EnsurePackage("RColorBrewer")

#Machine Learning
EnsurePackage("FactoMineR")
EnsurePackage("caret")
# EnsurePackage("keras")
EnsurePackage("rpart.plot")  # Decision Tree
EnsurePackage("randomForest") # Random Forest
EnsurePackage("class") # kNN
EnsurePackage("e1071")

cat("All Pacakges loaded")
```

```{r setup, include=FALSE,cache=FALSE}
## setting working directory
opts_knit$set(root.dir = "/Users/sathishrajendiran/Documents/R/CNN/data")
```

```{r }
cat("current working directory is",getwd())
```

```{r}
# image size to scale down to (original size varied)
width <- 28
height <- 28

## pbapply is a library to add progress bar *apply functions
## pblapply will replace lapply

extract_feature <- function(dir_path, width, height, label_id, add_label = TRUE) {
  img_size <- width*height
  ## List images in path
  images_names <- list.files(dir_path)
  if (add_label) {
    ## Set label
    label <- label_id
  }
  print(paste("Start processing", length(images_names), "images"))
  ## This function will resize an image, turn it into greyscale
  feature_list <- pblapply(images_names, function(imgname) {
    ## Read image
    img <- readImage(file.path(dir_path, imgname))
    ## Resize image
    img_resized <- resize(img, w = width, h = height)
    ## Set to grayscale
    grayimg <- channel(img_resized, "gray")
    ## Get the image as a matrix
    img_matrix <- grayimg@.Data
    ## Coerce to a vector
    img_vector <- as.vector(t(img_matrix))
    return(img_vector)
  })
  ## bind the list of vector into matrix
  feature_matrix <- do.call(rbind, feature_list)
  feature_matrix <- as.data.frame(feature_matrix)
  ## Set names
  names(feature_matrix) <- paste0("pixel", c(1:img_size))
  if (add_label) {
    ## Add label
    feature_matrix <- cbind(label = label, feature_matrix)
  }
  return(feature_matrix)
}
```


```{r}
## Task 2: Import myImages_20409 ##
image_20409_dir <- "/Users/sathishrajendiran/Documents/R/CNN/data/train/20409"
image_20409_dir
#use EBImage to resize the images to 28×28 and turn them into greyscale
train_data_20409 <- extract_feature(dir_path = image_20409_dir, width = width, height = height,label_id = 20409)
#Each image will be turned into a vector of length 784, with each element representing the value in a pixel
dim(train_data_20409) #[1] 879 785
```

```{r}
## Task 2: Import myImages_83144 ##
image_83144_dir <- "/Users/sathishrajendiran/Documents/R/CNN/data/train/83144"
image_83144_dir
#use EBImage to resize the images to 28×28 and turn them into greyscale
train_data_83144 <- extract_feature(dir_path = image_83144_dir, width = width, height = height,label_id = 83144)
#Each image will be turned into a vector of length 784, with each element representing the value in a pixel
dim(train_data_83144)
```



```{r}
## Task 2: Import myImages_113209 ##
image_113209_dir <- "/Users/sathishrajendiran/Documents/R/CNN/data/train/113209"
image_113209_dir
#use EBImage to resize the images to 28×28 and turn them into greyscale
train_data_113209 <- extract_feature(dir_path = image_113209_dir, width = width, height = height,label_id = 113209)
#Each image will be turned into a vector of length 784, with each element representing the value in a pixel
dim(train_data_113209)
```

```{r}
## Task 2: Import myImages_126637 ##
image_126637_dir <- "/Users/sathishrajendiran/Documents/R/CNN/data/train/126637"
image_126637_dir
#use EBImage to resize the images to 28×28 and turn them into greyscale
train_data_126637 <- extract_feature(dir_path = image_126637_dir, width = width, height = height,label_id = 126637)
#Each image will be turned into a vector of length 784, with each element representing the value in a pixel
dim(train_data_126637)
# head(train_data_126637)
```


```{r}
#Each image will be turned into a vector of length 784, with each element representing the value in a pixel
dim(train_data_20409) #[1] 879 785
dim(train_data_83144) #[1] 870 785
dim(train_data_113209) #[1] 567 785
dim(train_data_126637) #[1] 1115  785

## Bind rows in a single dataset
trainImagesAllDF <- rbind(train_data_20409, train_data_83144,train_data_113209,train_data_126637)
dim(trainImagesAllDF) #[1] 3431  785

#convert label column to factor
trainImagesAllDF$label <- as.factor(trainImagesAllDF$label)
dim(trainImagesAllDF)

str(trainImagesAllDF)
```


```{r}
# # compare number of images by labels
  x <- data.frame(table(trainImagesAllDF$label))
  colourCount = length(unique(trainImagesAllDF$label))
  getPalette = colorRampPalette(brewer.pal(5, "Set3"))
  
  gp_all <- ggplot(x, aes(x=Var1, y=Freq)) +   geom_bar(position="dodge",stat = "identity", fill=getPalette(colourCount))
  gp_all <- gp_all + ggtitle("Images by labels") + xlab("Labels") + ylab("Number of Images")
  gp_all <- gp_all +  theme(plot.title = element_text(hjust = 0.5),axis.title = element_text(),panel.background = element_blank())
  gp_all
```

```{r}
# 1. Training Set Preparation
  set.seed(234)  
  # lets split the Training data set int into training and test datasets.creates a value for dividing the data into train and test.
  # In this case the value is defined as   80% of the number of rows in the dataset
  sample_size = floor(0.80*nrow(trainImagesAllDF)) 

  # set seed to ensure you always have same random numbers generated #324 has 100% training accuracy 
  train_index = sample(seq_len(nrow(trainImagesAllDF)),size = sample_size)

  train_set =trainImagesAllDF[train_index,] #creates the training dataset with row numbers stored in train_index

  test_set=trainImagesAllDF[-train_index,]  # creates the test dataset excluding the row numbers mentioned in train_index

  cat("\nImages by Labels:")
  table(trainImagesAllDF$label)
  cat("\nTrain_set - Images by Labels:")
  table(train_set$label)
  cat("\nTest_set - Images by Labels:")
  table(test_set$label)
  
  dim(trainImagesAllDF)
  dim(train_set)
  dim(test_set)
  
```


```{r}
# # compare number of images by labels from train_set
# str(train_set)

  t_train <- data.frame(table(train_set$label))
  t_train
  colourCount = length(unique(train_set$label))
  getPalette = colorRampPalette(brewer.pal(5, "Set3"))

  gp_train_set <- ggplot(t_train, aes(x=Var1, y=Freq)) +   geom_bar(position="dodge",stat = "identity", fill=getPalette(colourCount))
  gp_train_set <- gp_train_set + ggtitle("Train data images by labels ") + xlab("Labels") + ylab("Number of Images")
  gp_train_set <- gp_train_set +  theme(plot.title = element_text(hjust = 0.5),axis.title = element_text(),panel.background = element_blank())
  gp_train_set
```


```{r}

# # compare number of images by labels from test_set
# str(test_set)

  t_test <- data.frame(table(test_set$label))
  colourCount = length(unique(test_set$label))
  getPalette = colorRampPalette(brewer.pal(5, "Set3"))

  gp_test_set <- ggplot(t_test, aes(x=Var1, y=Freq)) +   geom_bar(position="dodge",stat = "identity", fill=getPalette(colourCount))
  gp_test_set <- gp_test_set + ggtitle("Test data images by labels") + xlab("Labels") + ylab("Number of Images")
  gp_test_set <- gp_test_set +  theme(plot.title = element_text(hjust = 0.5),axis.title = element_text(),panel.background = element_blank())
  gp_test_set
  # t_test
```

```{r}
# Section 2: Build and tune decision tree models
    # grow tree
  rtree <- rpart(label~. ,data=train_set, method='class', cp=0,minsplit = 1, maxdepth = 5)
  #summarize rtree values
  summary(rtree)
  plotcp(rtree) # plot cross-validation results
  printcp(rtree) # plot cross-validation results
  # Plot tree | lets Plot decision trees
  rpart.plot(rtree,main="Image Classification Tree", extra= 102) # plot decision tree
  rsq.rpart(rtree) # plot approximate R-squared and relative error for different splits (2 plots)

```


```{r}
# Section 3: Prediction | Test Phase

  # cat("\nImages by Labels:")
  # table(trainImagesAllDF$label)
  # cat("\nTrain_set - Images by Labels:")
  # table(train_set$label)
  # cat("\nTest_set - Images by Labels:")
  # table(test_set$label)
  
  # dim(train_set)
  # dim(test_set)

  # Section 3: Prediction | Test Phase
  predict_unseen <- predict(rtree, test_set, type = 'class')
  # predict_unseen
  table_test_matrix <- table(test_set$label, predict_unseen)
  cat("\n\nPrediction results : Confusion Matrix \n\n")
  # table_mat
  confusionMatrix(table_test_matrix)
  
```


```{r}
# Random Forest prediction of Digit Tokenizer data

  # View(fedPapersDF1)
  cat("\n All Images by Labels:")
  table(test_set$label)
    
  fit <- randomForest(y=test_set$label, x=test_set[2:ncol(test_set)], data=test_set, ntree=50
                      , keep.forest=FALSE, importance=TRUE)
  
  print(fit) # view results
  importance(fit) # importance of each predictor
  
  rf_importance <- data.frame(importance(fit)) # importance of each predictor
  rf_importance
  
```

```{r}
#Principal Component Analysis
pca_digits <- PCA(t(select(trainImagesAllDF,-label)),ncp=30)
```

```{r}

imagePCATrainDF <- data.frame(trainImagesAllDF$label,pca_digits$var$coord)
dim(imagePCATrainDF) #[1] 3431   31

str(imagePCATrainDF)
# head(imagePCATrainDF)
```

```{r}
#Rename Columns 
colnames(imagePCATrainDF) <- c("label","dim1","dim2","dim3","dim4","dim5","dim6","dim7","dim8","dim9","dim10","dim11","dim12","dim13"
                                ,"dim14","dim15","dim16","dim17","dim18","dim19","dim20","dim21","dim22","dim23","dim24","dim25"
                                ,"dim26","dim27","dim28","dim29","dim30")
str(imagePCATrainDF)
dim(imagePCATrainDF)
# imagePCATrainDF
```

```{r}
# Prepare train and test data from the revised PCA Train dataset.
  sample_size = floor(0.80*nrow(imagePCATrainDF)) 
  # set seed to ensure you always have same random numbers generated #324 has 100% training accuracy 
  train_index = sample(seq_len(nrow(imagePCATrainDF)),size = sample_size)
  # train_index
  pca_train_set =imagePCATrainDF[train_index,] #creates the training dataset with row numbers stored in train_index
  # pca_train_set
  # # table(train_data$author)
  pca_test_set=imagePCATrainDF[-train_index,]  # creates the test dataset excluding the row numbers mentioned in train_index
  # # table(test_data$author)
  cat("\nImages by Labels:")
  table(imagePCATrainDF$label)
  cat("\nTrain_set - Images by Labels:")
  table(pca_train_set$label)
  cat("\nTest_set - Images by Labels:")
  table(pca_test_set$label)
```

```{r}
# # compare number of images by labels
  pca_x <- data.frame(table(imagePCATrainDF$label))
  colourCount = length(unique(imagePCATrainDF$label))
  getPalette = colorRampPalette(brewer.pal(9, "Set2"))
  pca_x
  pca_gp_all <- ggplot(pca_x, aes(x=Var1, y=Freq)) +   geom_bar(position="dodge",stat = "identity", fill=getPalette(colourCount))
  pca_gp_all <- pca_gp_all + ggtitle("Sample Images By Labels ") + xlab("Labels") + ylab("Number of Images")
  pca_gp_all <- pca_gp_all +  theme(plot.title = element_text(hjust = 0.5),axis.title = element_text(),panel.background = element_blank())
  pca_gp_all
  
```

```{r}
# # compare number of images by labels
  pca_tn_set <- data.frame(table(pca_train_set$label))
  colourCount = length(unique(pca_train_set$label))
  getPalette = colorRampPalette(brewer.pal(5, "Set3"))
  
  pca_gp_tnset <- ggplot(pca_tn_set, aes(x=Var1, y=Freq)) +   geom_bar(position="dodge",stat = "identity", fill=getPalette(colourCount))
  pca_gp_tnset <- pca_gp_tnset + ggtitle("Train Set Imagesby Label") + xlab("Labels") + ylab("Number of Images")
  pca_gp_tnset <- pca_gp_tnset +  theme(plot.title = element_text(hjust = 0.5),axis.title = element_text(),panel.background = element_blank())
  pca_gp_tnset
```

```{r}
# # compare number of images by labels
  pca_tst_set <- data.frame(table(pca_test_set$label))
  colourCount = length(unique(pca_test_set$label))
  getPalette = colorRampPalette(brewer.pal(5, "Set3"))
  
  pca_gp_tst_set <- ggplot(pca_tst_set, aes(x=Var1, y=Freq)) +   geom_bar(position="dodge",stat = "identity", fill=getPalette(colourCount))
  pca_gp_tst_set <- pca_gp_tst_set + ggtitle("Test data images by labels") + xlab("Labels") + ylab("Number of Images")
  pca_gp_tst_set <- pca_gp_tst_set +  theme(plot.title = element_text(hjust = 0.5),axis.title = element_text(),panel.background = element_blank())
  pca_gp_tst_set
```

```{r}
# Section 2: Build and tune decision tree models with PCA
  # grow tree
  pca_rtree <- rpart(label~. ,data=pca_train_set, method='class', cp=0,minsplit = 1, maxdepth = 5)
  #summarize rtree values
  summary(pca_rtree)
  plotcp(pca_rtree) # plot cross-validation results
  printcp(pca_rtree) # plot cross-validation results
  # Plot tree | lets Plot decision trees
  rpart.plot(pca_rtree,main="Classification Tree for Images after PCA", extra= 102) # plot decision tree
  rsq.rpart(pca_rtree) # plot approximate R-squared and relative error for different splits (2 plots)
```

```{r}
# Section 3: Prediction | Test Phase

  # cat("\nImages by Labels:")
  # table(imagePCATrainDF$label)
  # cat("\nTrain_set - Images by Labels:")
  # table(pca_train_set$label)
  # cat("\nTest_set - Images by Labels:")
  # table(pca_test_set$label)
  # Section 3: Prediction | Test Phase
  pca_predict_unseen <- predict(pca_rtree, pca_test_set, type = 'class')
  # predict_unseen
  pca_table_test_matrix <- table(pca_test_set$label, pca_predict_unseen)
  cat("\n\nPrediction results : Confusion Matrix \n\n")
  # table_mat
  confusionMatrix(pca_table_test_matrix)
```

```{r}
# kNN - Training the KNN model
pca_training_kNN <- trainControl(method = "repeatedcv", number = 10, repeats = 3)
set.seed(3033)
pca_knn_fit <- train(label ~ ., data = pca_train_set, method = "knn", trControl=pca_training_kNN
                     , preProcess = c("center", "scale")
                     , tuneLength = 10)
pca_knn_fit
```

```{r}
#kNN - Prediction
pca_knn_test_pred <- predict(pca_knn_fit, newdata =pca_test_set)
View(data.frame(pca_knn_test_pred, pca_test_set$label))
summary(pca_knn_test_pred)
```

```{r}
#SVM - Building the Model | Linear Kernel

pca_svm_model <- svm(label ~ ., data = pca_train_set, kernel= 'linear', cost =100, scale = FALSE
                     , Probability = TRUE, Cross = 3,type = 'C')
summary(pca_svm_model)
```

```{r}
#SVM - Prediction | Linear Kernel
pca_svmPred <- predict(pca_svm_model, pca_test_set, type = 'prob')
pca_svmPred <- as.data.frame(pca_svmPred)
colnames(pca_svmPred) <- 'results'


svmResults <- pca_test_set %>% select(label) %>% bind_cols(pca_svmPred) %>% 
  mutate(real = factor(as.character(str_remove(label, 'V')))
         , prediction = factor(as.character(str_remove(results, 'V'))))

confusionMatrix(svmResults$real, svmResults$prediction) #54.88%
```

```{r}
#SVM - Building the Model | Radial Kernel | cost 10
pca_svm_model_radial <- svm(label ~ ., data = pca_train_set, kernel= 'radial', cost =100
                            , scale = FALSE, Probability = TRUE, Cross = 3,type = 'C')
summary(pca_svm_model_radial)

#SVM - Prediction
pca_svmPred_radial <- predict(pca_svm_model_radial, pca_test_set, type = 'prob')

#Plot the results
pca_svmPred_radial <- as.data.frame(pca_svmPred_radial)
colnames(pca_svmPred_radial) <- 'results'
svmResults_radial <- pca_test_set %>% select(label) %>% bind_cols(pca_svmPred_radial) %>% mutate(real = factor(as.character(str_remove(label, 'V'))) , prediction = factor(as.character(str_remove(results, 'V'))))
#Build Confusion matrix
pca_SVM_radialMatrix <- confusionMatrix(svmResults_radial$real
                                        , svmResults_radial$prediction) #94.48 %
pca_SVM_radialMatrix
pca_SVM_radialMatrix$overall
```

```{r}
 #SVM - Building the Model | Ploynomial Kernel | cost 100
pca_svm_model_poly <- svm(label ~ ., data = pca_train_set, kernel= 'polynomial', cost =100
                          , scale = FALSE, Probability = TRUE, Cross = 3,type = 'C')
summary(pca_svm_model_poly)

#SVM - Prediction
pca_svmPred_poly <- predict(pca_svm_model_poly, pca_test_set, type = 'prob')

summary(pca_svmPred_poly)

#Plot the results
pca_svmPred_poly <- as.data.frame(pca_svmPred_poly)
colnames(pca_svmPred_poly) <- 'results'
svmResults_poly <- pca_test_set %>% select(label) %>% bind_cols(pca_svmPred_poly) %>% mutate(real = factor(as.character(str_remove(label, 'V')))
                                                                                                 , prediction = factor(as.character(str_remove(results, 'V'))))
#Build Confusion matrix
pca_SVM_polyMatrix <- confusionMatrix(svmResults_poly$real, svmResults_poly$prediction) #94.48 %
pca_SVM_polyMatrix
pca_SVM_polyMatrix$overall
```

```{r}
# Random Forest prediction of PCA data
  rf_fit <- randomForest(y=pca_train_set$label, x=pca_train_set[2:ncol(pca_train_set)], data=pca_train_set, ntree=100 , keep.forest=FALSE, importance=TRUE)
  
  print(rf_fit) # view results
  importance(rf_fit) # importance of each predictor
  
  rf_importance <- data.frame(importance(rf_fit)) # importance of each predictor
  rf_importance
```

## Now lets rebuild the models using Binary variables. In this case, Labels 20409 and 83144 seems evenly distributed. Lets, go with these categorical values.
```{r}
## Bind rows in a single dataset
trainImages2DF <- rbind(train_data_20409, train_data_83144)
dim(trainImages2DF) #[1] 3431  785

#convert label column to factor
trainImages2DF$label <- as.factor(trainImages2DF$label)
dim(trainImages2DF)
str(trainImages2DF)

```

```{r}
#Principal Component Analysis
pca_2digits <- PCA(t(select(trainImages2DF,-label)),ncp=30)
```



```{r}
#Prepare dataframe from PCA Output
trainPCAImages2DF <- data.frame(trainImages2DF$label,pca_2digits$var$coord)
dim(trainPCAImages2DF) #[1] 3431   31

str(trainPCAImages2DF)
# head(imagePCATrainDF)
```

```{r}
#Rename Columns 
colnames(trainPCAImages2DF) <- c("label","dim1","dim2","dim3","dim4","dim5","dim6","dim7","dim8","dim9","dim10","dim11","dim12","dim13"
                                ,"dim14","dim15","dim16","dim17","dim18","dim19","dim20","dim21","dim22","dim23","dim24","dim25"
                                ,"dim26","dim27","dim28","dim29","dim30")
str(trainPCAImages2DF)
dim(trainPCAImages2DF)
# imagePCATrainDF
```

```{r}
# Prepare train and test data from the revised PCA Train dataset.
  sample_size = floor(0.80*nrow(trainPCAImages2DF)) 
  # set seed to ensure you always have same random numbers generated #324 has 100% training accuracy 
  train_index = sample(seq_len(nrow(trainPCAImages2DF)),size = sample_size)
  # train_index
  pca_train_set_binary =trainPCAImages2DF[train_index,] #creates the training dataset with row numbers stored in train_index
  # pca_train_set
  # # table(train_data$author)
  pca_test_set_binary=trainPCAImages2DF[-train_index,]  # creates the test dataset excluding the row numbers mentioned in train_index
  # # table(test_data$author)
  cat("\nImages by Labels:")
  table(trainPCAImages2DF$label)
  cat("\nTrain_set - Images by Labels:")
  table(pca_train_set_binary$label)
  cat("\nTest_set - Images by Labels:")
  table(pca_test_set_binary$label)
```
  
  
  
```{r}
head(trainPCAImages2DF)
imagePCATrain2DF <- subset(trainPCAImages2DF, select = c(1, 2, 5))
imagePCATrain2DF
# # compare number of images by labels
  pca_2x <- data.frame(table(trainPCAImages2DF$label))
  colourCount = length(unique(trainPCAImages2DF$label))
  getPalette = colorRampPalette(brewer.pal(5, "Set2"))
  pca_2x
  pca_gp_2x <- ggplot(pca_2x, aes(x=Var1, y=Freq)) +   geom_bar(position="dodge",stat = "identity", fill=getPalette(colourCount))
  pca_gp_2x <- pca_gp_2x + ggtitle("Sample Images By Labels ") + xlab("Labels") + ylab("Number of Images")
  pca_gp_2x <- pca_gp_2x +  theme(plot.title = element_text(hjust = 0.5),axis.title = element_text(),panel.background = element_blank())
  pca_gp_2x
  
```

```{r}
# # compare number of images by labels
  pca_tn_set_binary <- data.frame(table(pca_train_set_binary$label))
  colourCount = length(unique(pca_train_set_binary$label))
  getPalette = colorRampPalette(brewer.pal(5, "Set3"))
  
  pca_gp_tnset_binary <- ggplot(pca_tn_set_binary, aes(x=Var1, y=Freq)) +   geom_bar(position="dodge",stat = "identity", fill=getPalette(colourCount))
  pca_gp_tnset_binary <- pca_gp_tnset_binary + ggtitle("Train Set Imagesby Label") + xlab("Labels") + ylab("Number of Images")
  pca_gp_tnset_binary <- pca_gp_tnset_binary +  theme(plot.title = element_text(hjust = 0.5),axis.title = element_text(),panel.background = element_blank())
  pca_gp_tnset_binary
```

```{r}
# # compare number of images by labels
  pca_tst_set_binary <- data.frame(table(pca_test_set_binary$label))
  colourCount = length(unique(pca_test_set_binary$label))
  getPalette = colorRampPalette(brewer.pal(5, "Set3"))
  
  pca_gp_tst_set_binary <- ggplot(pca_tst_set_binary, aes(x=Var1, y=Freq)) +   geom_bar(position="dodge",stat = "identity", fill=getPalette(colourCount))
  pca_gp_tst_set_binary <- pca_gp_tst_set_binary + ggtitle("Test data images by labels") + xlab("Labels") + ylab("Number of Images")
  pca_gp_tst_set_binary <- pca_gp_tst_set_binary +  theme(plot.title = element_text(hjust = 0.5),axis.title = element_text(),panel.background = element_blank())
  pca_gp_tst_set_binary
```

```{r}
# Section 2: Build and tune decision tree models with PCA
  # grow tree
  pca_rtree_binary <- rpart(label~. ,data=pca_train_set_binary, method='class', cp=0,minsplit = 1, maxdepth = 5)
  #summarize rtree values
  summary(pca_rtree_binary)
  plotcp(pca_rtree_binary) # plot cross-validation results
  printcp(pca_rtree_binary) # plot cross-validation results
  # Plot tree | lets Plot decision trees
  rpart.plot(pca_rtree_binary,main="Classification Tree for Images after PCA", extra= 102) # plot decision tree
  rsq.rpart(pca_rtree_binary) # plot approximate R-squared and relative error for different splits (2 plots)
```

```{r}
# Section 3: Prediction | Test Phase
  pca_predict_unseen_binary <- predict(pca_rtree_binary, pca_test_set_binary, type = 'class')
  # predict_unseen
  pca_table_test_matrix_binary <- table(pca_test_set_binary$label, pca_predict_unseen_binary)
  cat("\n\nPrediction results : Confusion Matrix \n\n")
  # table_mat
  confusionMatrix(pca_table_test_matrix_binary)
```

```{r}
# kNN - Training the KNN model
pca_training_kNN_binary <- trainControl(method = "repeatedcv", number = 10, repeats = 3)
set.seed(3033)
pca_knn_fit_binary <- train(label ~ ., data = pca_train_set_binary, method = "knn", trControl=pca_training_kNN_binary
                     , preProcess = c("center", "scale")
                     , tuneLength = 10)
pca_knn_fit_binary
```

```{r}
#kNN - Prediction
pca_knn_test_pred_binary <- predict(pca_knn_fit_binary, newdata =pca_test_set_binary)
View(data.frame(pca_knn_test_pred_binary, pca_test_set_binary$label))
summary(pca_knn_test_pred_binary)
```

```{r}
#SVM - Building the Model | Linear Kernel | Binary Classifier

pca_svm_model_binary <- svm(label ~ ., data = pca_train_set_binary, kernel= 'linear'
                            , cost =100, scale = FALSE, Probability = TRUE
                            , Cross = 3,type = 'C')
summary(pca_svm_model_binary)
```

```{r}
#SVM - Prediction | Linear Kernel | Binary Classifier
pca_svmPred_binary <- predict(pca_svm_model_binary, pca_test_set_binary, type = 'prob')
pca_svmPred_binary <- as.data.frame(pca_svmPred_binary)
colnames(pca_svmPred_binary) <- 'results'


svmResults_binary <- pca_test_set_binary %>% select(label) %>% bind_cols(pca_svmPred_binary)   %>% mutate(real = factor(as.character(str_remove(label, 'V')))
         , prediction = factor(as.character(str_remove(results, 'V'))))

confusionMatrix(svmResults_binary$real, svmResults_binary$prediction) #73.71%
```

```{r}
#SVM - Building the Model | Radial Kernel | cost 100 | Binary Classifier
pca_svm_model_radial_binary <- svm(label ~ ., data = pca_train_set_binary, kernel= 'radial', cost =100
                            , scale = FALSE, Probability = TRUE, Cross = 3,type = 'C')
summary(pca_svm_model_radial_binary)

#SVM - Prediction
pca_svmPred_radial_binary <- predict(pca_svm_model_radial_binary, pca_test_set_binary, type = 'prob')

#Plot the results
pca_svmPred_radial_binary <- as.data.frame(pca_svmPred_radial_binary)
colnames(pca_svmPred_radial_binary) <- 'results'
svmResults_radial_binary <- pca_test_set_binary %>% select(label) %>% bind_cols(pca_svmPred_radial_binary) %>% mutate(real = factor(as.character(str_remove(label, 'V'))) , prediction = factor(as.character(str_remove(results, 'V'))))
#Build Confusion matrix
pca_SVM_radialMatrix_binary <- confusionMatrix(svmResults_radial_binary$real
                                        , svmResults_radial_binary$prediction) #76.57%
pca_SVM_radialMatrix_binary
pca_SVM_radialMatrix_binary$overall
```

```{r}
 #SVM - Building the Model | Ploynomial Kernel | cost 100 | Binary Classifier
pca_svm_model_poly_binary <- svm(label ~ ., data = pca_train_set_binary, kernel= 'polynomial', cost =100
                          , scale = FALSE, Probability = TRUE, Cross = 3,type = 'C')
summary(pca_svm_model_poly_binary)

#SVM - Prediction
pca_svmPred_poly_binary <- predict(pca_svm_model_poly_binary, pca_test_set_binary, type = 'prob')

summary(pca_svmPred_poly_binary)

#Plot the results
pca_svmPred_poly_binary <- as.data.frame(pca_svmPred_poly_binary)
colnames(pca_svmPred_poly_binary) <- 'results'
svmResults_poly_binary <- pca_test_set_binary %>% select(label) %>% bind_cols(pca_svmPred_poly_binary) %>% mutate(real = factor(as.character(str_remove(label, 'V')))
                                                                                                 , prediction = factor(as.character(str_remove(results, 'V'))))
#Build Confusion matrix
pca_SVM_polyMatrix_binary <- confusionMatrix(svmResults_poly_binary$real, svmResults_poly_binary$prediction) #94.48 %
pca_SVM_polyMatrix_binary
pca_SVM_polyMatrix_binary$overall
```

```{r}
# Random Forest prediction of PCA data | Binary Classifier
  rf_fit_binary <- randomForest(y=pca_train_set_binary$label, x=pca_train_set_binary[2:ncol(pca_train_set_binary)], data=pca_train_set_binary, ntree=100 , keep.forest=FALSE, importance=TRUE)
  
  print(rf_fit_binary) # view results
  importance(rf_fit_binary) # importance of each predictor
  
  rf_importance_binary <- data.frame(importance(rf_fit_binary)) # importance of each predictor
  rf_importance_binary
```

```{r}
# images_20409 <-list.files(path=image_20409_dir, pattern=".jpg",all.files=T)
# myImages_20409 = lapply(images_20409, readImage)

# n_files <- length(images_20409)
# tmpfiles <- .3 * n_files
# 
# for (i in 1:tmpfiles) {
#   myImages_20409[i] = lapply(images_20409[i], readImage)
# }

# length(myImages_20409)
#structure of the list
# str(myImages_20409)
#image as in 
# print(myImages[1])
```


